Investigating the alternative prey hypothesis with the POMP framework

### 2024-04-18 workspace/sizhuang.Rmd
sizhuang EDA

```{r data_table, echo=FALSE}
data <- read_excel("./data/bird_data.xlsx")
data |>
  rename(peak_rodent_year = "peak rodent year") -> data

knitr::kable(data[1:5, 1:6], caption = "Data")
```

```{r, echo=FALSE, message=FALSE}
# Most basic bubble plot
p1 <- ggplot(data, aes(x=year, y=CPUE)) +
  geom_line(color="steelblue") + 
  geom_point() + 
  geom_rect(
    data = subset(data, peak_rodent_year == 'yes'),
            aes(xmin = year - 0.5, xmax = year + 0.5, ymin = -Inf, ymax = Inf),
            fill = "blue", alpha = 0.3
    ) + 
  theme_light() +
  xlab("") + 
  stat_smooth(
    color = "#FC4E07",
    method = "loess"
  )

p2 <- ggplot(data, aes(x=year, y=log.CPUE)) + 
  geom_line(color="steelblue") + 
  geom_point() + 
  geom_rect(
    data = subset(data, peak_rodent_year == 'yes'),
            aes(xmin = year - 0.5, xmax = year + 0.5, ymin = -Inf, ymax = Inf),
            fill = "blue", alpha = 0.3
    ) + 
  theme_light() + 
  xlab("Year") + 
  stat_smooth(
    color = "#FC4E07",
    method = "loess"
  )

@@

```{r ACF, echo=FALSE}
acf(data$log.CPUE, lag = 100, main = "ACF log.CPUE")
```

```{r PACF, echo=FALSE}
pacf(data$log.CPUE, lag = 100, main = "PACF log.CPUE")
```

>>>>

```{r data_table, echo=FALSE}
data <- read_excel("../data/bird_data.xlsx")
data |>
  rename(peak_rodent_year = "peak rodent year") -> data

data[data == "NA"] <- NA
clean_data <- na.omit(data)

# knitr::kable(clean_data[1:5, 1:6], caption = "Cleaned Data")
```

```{r}
columns_to_convert <- c("growth rate CPUE", "temp.PREINC", "temp.INC", "temp.HATCH", "temp.BROOD", "temp.WINTER", "temp.SPRING", "prec.PREINC", "prec.INC", "prec.HATCH", "prec.BROOD")
# Convert these columns to numeric
clean_data[columns_to_convert] <- lapply(clean_data[columns_to_convert], as.numeric)

# Check for any remaining NAs
# sapply(clean_data[columns_to_convert], function(x) sum(is.na(x)))
```

```{r}
# Summary statistics for all numerical variables
summary(clean_data)

```{r, echo=FALSE, message=FALSE}
# Most basic bubble plot
p1 <- ggplot(clean_data, aes(x=year, y=CPUE)) +
  geom_line(color="steelblue") + 
  geom_point() + 
  geom_rect(
    data = subset(clean_data, peak_rodent_year == 'yes'),
            aes(xmin = year - 0.5, xmax = year + 0.5, ymin = -Inf, ymax = Inf),
            fill = "blue", alpha = 0.3
    ) + 
  theme_light() +
  xlab("") + 
  stat_smooth(
    color = "#FC4E07",
    method = "loess"
  )

p2 <- ggplot(clean_data, aes(x=year, y=log.CPUE)) + 
  geom_line(color="steelblue") + 
  geom_point() + 
  geom_rect(
    data = subset(clean_data, peak_rodent_year == 'yes'),
            aes(xmin = year - 0.5, xmax = year + 0.5, ymin = -Inf, ymax = Inf),
            fill = "blue", alpha = 0.3
    ) + 
  theme_light() + 
  xlab("Year") + 
  stat_smooth(
    color = "#FC4E07",
    method = "loess"
  )

@@

```{r ACF, echo=FALSE}
acf(clean_data$log.CPUE, lag = 100, main = "ACF log.CPUE")
```

```{r PACF, echo=FALSE}
pacf(clean_data$log.CPUE, lag = 100, main = "PACF log.CPUE")
```

```{r}
# Box plot of CPUE for each Peak Rodent Year
ggplot(data, aes(x = peak_rodent_year, y = CPUE, fill = peak_rodent_year)) +
  geom_boxplot() +
  labs(title = "CPUE by Peak Rodent Year", x = "Peak Rodent Year", y = "CPUE")
```
